---
title: "Example 2.4.1 - Fitting a path model with observed variables only, extracting model parameter matrices and tables in APA format"
output:
  pdf_document: default
  word_document: default
---

# Fitting a path model with observed variables only

First, we execute the code from the book. We load the package:

```{r, message = FALSE, warning = FALSE}
library("lavaan")
library("knitr")
```

Next, we load the input data (a covariance matrix, in this case):

```{r}
beaujean.cov <- lav_matrix_lower2full(c(648.07, 30.05, 8.64, 140.18, 25.57, 233.21))
colnames(beaujean.cov) <- rownames(beaujean.cov) <- c("salary", "school", "iq")
```

Next, we specify the model:

```{r}
beaujean.model <- '
  salary ~ a*school + c*iq
  iq ~ b*school # this was reversed in first printing of the book 
  ind:= b*c 
'
```

Finally, we fit the model to the data and inspect the result:

```{r}
beaujean.fit <- sem(beaujean.model, sample.cov=beaujean.cov, sample.nobs=300)
summary(beaujean.fit, standardized = TRUE, fit.measures = TRUE, rsquare = TRUE, 
        modindices = TRUE)
```


# Extracting matrices with parameter estimates

The parameter estimates (or coefficients) can be obtained from the fitted model as follows:

```{r}
coefs <- inspect(beaujean.fit, "coef")
coefs$beta
coefs$psi
```

Here we see that $\psi$ is a symmetric matrix, giving the (co)variances; $\beta$ is a non-symmetric matrix, giving the regression coefficients. 


# Obtaining tables for a paper

To copy the results to a paper, we use function `kable` from package `knitr`:

```{r}
kable(parameterEstimates(beaujean.fit), digits = 3)
```

Note that you can also compile Markdown documents (.Rmd files) as Word documents. As you can see in the github repository, the documents for this example are available as an R Markdown file, which allows you to combine text, code and results. The code can be run and the document compiled in R Studio by clicking the "Knit" button. 

Check out the Markdown file and compare to the compiled .docx and .pdf files to see how the look of the final document can be controlled.

The table still needs some manual adjustments, of course (e.g., $p$-values should never be written as 0.000, but as $< .001). Note that you can use function `kable` from package `kableExtra` to have more control over how the final tables look. This can be very helpful for publications (but outside the scope of this course).

```{r, echo=FALSE}
## If I want to hide a piece of code, I use "echo = FALSE" as a chunk options
## By default, "echo=TRUE", so all R code will be printed to the document
```

```{r, echo=FALSE, results='hide'}
## If I want to run a piece of code, but not show the output, I use "results = 'hide'"
```

```{r, echo=FALSE, eval=FALSE}
## If I want to show a piece of code in the document, but not run it, I use "eval=FALSE"
```
